---
title: "Homework 5"
author: "Krishna Vemulapalli"
output: github_document
---

```{r setup}
library(tidyverse)
library(ggplot2)
library(knitr)
library(janitor)
library(purrr)
```

## Problem 1

Describing the raw dataset:     



Reading and cleaning the dataset. Followed by, creating a city_state variable and summarizing within cities. 
```{r}
homocide_df = 
  read.csv("data/homicide-data.csv") %>%
  mutate(
    city_state = str_c(city, state, sep = ", "), #creating the city_state variable#
    homocide_status = case_when(
     disposition == "Open/No arrest" ~ "unsolved",
     disposition == "Closed without arrest" ~ "unsolved",
     disposition == "Closed by arrest" ~ "solved"
    ) #creating the homocide status variable#
  ) %>%
  clean_names()

summary_df = 
  homocide_df %>%
  filter(city_state != "Tulsa, AL") %>% #removed the "Tulsa, AL" obs because it has only one unsolved homocide. This indicates a data entry mistake#
  select(city_state, homocide_status) %>% #keeping only city_state and homocide_status variable#
  group_by(city_state) %>%
  summarise(
    total_no_homocides = n(),
    no_unsolved_homocides = sum(homocide_status == "unsolved")
  ) #calculating the total homocides and number of unsolved homocides by city_state#

summary_df %>%
  kable()
  
```

Estimating the proportion of homicides that are unsolved in Baltimore, MD
```{r}
baltimore_MD_prop_test = 
  prop.test(
  summary_df %>% filter(city_state == "Baltimore, MD") %>% pull(no_unsolved_homocides),
  summary_df %>% filter(city_state == "Baltimore, MD") %>% pull(total_no_homocides))

#Applying the broom::tidy to convert the output into a tibble#
baltimore_MD_prop_test %>% 
  broom::tidy() %>%
  select(estimate, conf.low, conf.high) #selecting only the estimate and confidence intervals
```

Estimating the proportion of homicides that are unsolved and creating estimate and confidence intervals for each city_state
```{r}
city_state_prop_test =
  summary_df %>%
  mutate(
    prop_test = map2(.x = no_unsolved_homocides, .y = total_no_homocides, ~prop.test(.x, .y)),
    prop_test_tidy = map(.x = prop_test, ~broom::tidy(.x))
  ) %>%
  select(-prop_test) %>% #dropping prop_test variable#
  unnest(prop_test_tidy) %>% #unnesting prop_test_tidy variable#
  select(city_state, estimate, conf.low, conf.high) %>% #subsetting only the city_state, estimate and confidence intervals#
  arrange(desc(estimate)) #Organizing cities according to the proportion of unsolved homicides#
```

Creating a plot that shows the estimates and CIs for each city
```{r}
city_state_prop_test %>%
  mutate(city_state = fct_reorder(city_state, estimate)) %>%
  ggplot(aes(x = city_state, y = estimate)) +
  geom_point() +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high)) +
  theme(axis.text.x = element_text(angle = 90)) +
  ylab("Proportion of Unsolved Homicides") + 
  xlab("City, State")
```

